---
title: "AcquisitionCSCode"
author: "Robert Perez"
date: "6/3/2021"
output: html_document
---

### Importing the Data

```{r data1}
DJI <- read.csv("dow_jones_index.data") #25 of each stock in the dataset
```

### Cleaning Data

```{r data2}
DJI1 <- DJI

DJI1$quarter <- as.factor(DJI1$quarter)

DJI1$date <- as.Date(DJI1$date, "%m/%d/%Y")
DJI1$open = as.numeric(gsub("\\$", "", DJI1$open))
DJI1$high = as.numeric(gsub("\\$", "", DJI1$high))
DJI1$low = as.numeric(gsub("\\$", "", DJI1$low))
DJI1$close = as.numeric(gsub("\\$", "", DJI1$close))
DJI1$next_weeks_close = as.numeric(gsub("\\$", "", DJI1$next_weeks_close))
DJI1$next_weeks_open = as.numeric(gsub("\\$", "", DJI1$next_weeks_open))
DJI1$stock <- as.factor(DJI1$stock)
stocknames = levels(DJI1$stock)

```

```{r data3}

Week = strftime(DJI1$date, format = "%V")
DJI1 = cbind(Week, DJI1)
DJI1$Week <- as.numeric(DJI1$Week)
DJI1 <- DJI1[order(DJI1$stock),]
str(DJI1)
```

```{r data0}
DJI1 <- DJI1[-c(13:14,16:17)] #Not using "next" columns as predictors
```

### Creating lag plots

```{r}
CNamesFunction <- function(dataframe)
{
  ColumnNames = colnames(dataframe)
}
```

```{r}
LagPlotFunc <- function(column, lagnumber) 
{
  lag.plot(column, set.lags = 1:lagnumber)
}

```

```{r}
LagVars = CNamesFunction(Filter(is.numeric, DJI1)[,1:10]) # only first 10 columns, all other numeric variables are lag variables

for(i in LagVars)
{ 
data = DJI1[[i]]  
 LagPlotFunc(na.omit(data),4)
}
```


### Creating lag variables

```{r data4}
LagValFunction <-  function(variname) 
{
  print(lag(variname, 1))
}
```

```{r data5}
open.lag <- c()
high.lag <- c()
low.lag <- c()
close.lag <- c()
volume.lag <- c()
percent_change_price.lag <- c()
percent_change_volume_over_last_wk.lag <- c()
previous_weeks_volume.lag <- c()
percent_change_next_weeks_price.lag <- c()
```

```{r data6, results="hide"}
for(i in stocknames)
{ 
  x = filter(DJI1,stock == i)
  
  open.lag <- c(open.lag,LagValFunction(x$open))
  high.lag <- c(high.lag,LagValFunction(x$high))
  low.lag <- c(low.lag,LagValFunction(x$low))
  close.lag <- c(close.lag,LagValFunction(x$close))
  volume.lag <- c(volume.lag,LagValFunction(x$volume))
  percent_change_price.lag <- c(percent_change_price.lag,LagValFunction(x$percent_change_price))
  percent_change_volume_over_last_wk.lag <- c(percent_change_volume_over_last_wk.lag,LagValFunction(x$percent_change_volume_over_last_wk))
  previous_weeks_volume.lag <- c(previous_weeks_volume.lag,LagValFunction(x$previous_weeks_volume))
  percent_change_next_weeks_price.lag <- c(percent_change_next_weeks_price.lag,LagValFunction(x$percent_change_next_weeks_price))

}

DJI1 <- cbind(DJI1,open.lag, high.lag, low.lag, close.lag,volume.lag,percent_change_price.lag,percent_change_volume_over_last_wk.lag,previous_weeks_volume.lag,percent_change_next_weeks_price.lag)
```

### Splitting the Data

```{r data7}
DJI1 <- na.omit(DJI1)  #Removes the 1/7/2011 week from all stocks, because the there are missing percent change variables.  This is because they are the first variable in the time series.
head(DJI1)
```

```{r data8}
DJI1Training <- DJI1[ which(DJI1$quarter=='1'),]
DJI1Test <- DJI1[ which(DJI1$quarter=='2'),]
```

### Scatter Plots

```{r}
PlotAxis = CNamesFunction(DJI1)
```

```{r}
ScatterFunction <- function(column, var_name) 
{
  plot(DJI1$percent_change_next_weeks_price, column, main="Next Week Percent Change Scatterplot",
   xlab="Percent Change in Next Weeks Price ", ylab = var_name, pch=19)
}
```

```{r}
for(i in PlotAxis)
{ 
data = DJI1[[i]]  
 ScatterFunction(data,i)
}
```

```{r}
ScatterFunction <- function(response, column, xlabel, ylabel) 
{
  plot(response, column, main="Next Week Percent Change Scatterplot",
   xlab= xlabel , ylab = ylabel, pch=19)
  
  points(predict(lm(column ~ response)), col = "red", pch = 19)
}
  
  
for (i in stocknames)
{
  x = filter(DJI1, stock == i)
  
  ScatterFunction(x$Week, x$percent_change_next_weeks_price,  "Week", "Percent Change in Next Week's Price")
}
```


### Determing the formula for regression

```{r}
DJI2 <- na.omit(DJI1) #DJI2 omits NAs in DJI1 in order for "cor" function to generate correlation values for lag variables.
num_cols <- unlist((lapply(DJI2, is.numeric)))
CorValues <- data.frame(cor(DJI2[, na.omit(num_cols)]))
PerChangeNextCorVals <- CorValues[10]
PerChangeNextCorVals <- PerChangeNextCorVals %>% rownames_to_column('Variable') #preserves variable names for ease of access
PerChangeNextCorVals <- PerChangeNextCorVals[-c(10),] #removing the correlation between Next Weeks Price and itself
```

```{r}
AveragePerCorVal <- mean(PerChangeNextCorVals$percent_change_next_weeks_price)
AveragePerCorVal
```

```{r}
FormulaCorVals <- (PerChangeNextCorVals[PerChangeNextCorVals$percent_change_next_weeks_price > AveragePerCorVal,])

View(FormulaCorVals)

```

```{r}
(FormulaCorVals$Variable)
```

```{r}
LM1 = lm(percent_change_next_weeks_price ~ Week + open + high + low + close + percent_change_price + percent_change_volume_over_last_wk + open.lag + close.lag + high.lag + low.lag + percent_change_volume_over_last_wk.lag + percent_change_next_weeks_price.lag, data = DJI1) 
#car::vif(LM1) #<- cannot run VIF due to aliased coefficients
```

```{r}
ld.vars <- attributes(alias(LM1)$Complete)$dimnames[[1]]
print(ld.vars) # Outputs aliased coefficients
```

```{r}
LM2 = lm(percent_change_next_weeks_price ~ Week + open + high + low + close + percent_change_price + percent_change_volume_over_last_wk + open.lag + close.lag + high.lag + low.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM2)
```

```{r}
LM3 = lm(percent_change_next_weeks_price ~ Week + open + high + low + close + percent_change_price + percent_change_volume_over_last_wk + open.lag + high.lag + low.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM3)
```

```{r}
LM4 = lm(percent_change_next_weeks_price ~ Week + high + low + close + percent_change_price + percent_change_volume_over_last_wk + open.lag + high.lag + low.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM4)
```

```{r}
LM5 = lm(percent_change_next_weeks_price ~ Week + high + low + percent_change_price + percent_change_volume_over_last_wk + open.lag + high.lag + low.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM5)
```

```{r}
LM6 = lm(percent_change_next_weeks_price ~ Week + high + low + percent_change_price + percent_change_volume_over_last_wk + open.lag + low.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM6)
```

```{r}
LM7 = lm(percent_change_next_weeks_price ~ Week + high + low + percent_change_price + percent_change_volume_over_last_wk + open.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM7)
```

```{r}
LM8 = lm(percent_change_next_weeks_price ~ Week + low + percent_change_price + percent_change_volume_over_last_wk + open.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM8)
```

```{r}
LM9 = lm(percent_change_next_weeks_price ~ Week + percent_change_price + percent_change_volume_over_last_wk + open.lag + percent_change_volume_over_last_wk.lag, data = DJI1) 
car::vif(LM9)
```

```{r data9}
formulafinal = percent_change_next_weeks_price ~ Week + percent_change_price + percent_change_volume_over_last_wk + open.lag + percent_change_volume_over_last_wk.lag
```

### Creating the Linear Model and MSE

```{r}
StockFunctionL <- function (training, testing, formula)  {
  set.seed(111)

  LinearModel <- lm(formula, data = training)

  LinearPreds <- predict(LinearModel, newdata = testing)

  LinearSumm <- (summary(LinearModel))

  MSE.Linear <- (mean((testing$percent_change_next_weeks_price - LinearPreds)^2))

  print(MSE.Linear)
}
```

```{r}
MSE.L <- c()

for (i in stocknames)
{x = filter((DJI1Training), stock == i)

  y = filter(DJI1Test, stock == i)

  MSE.L <- c(MSE.L,StockFunctionL(x, y, formulafinal))
  
  
  
}
view(MSE.L)

df <- data.frame(stocknames, MSE.L)
```

### Creating the Tree Model and MSE

```{r}
StockFunctionTree <- function(training, testing, formula) {
  set.seed(111)
  
  TreeModel <- tree(formula, data = training)
 
  TreePreds <- predict(TreeModel, newdata = testing)
  
  MSE.Tree <- (mean((testing$percent_change_next_weeks_price - TreePreds)^2))

  print(MSE.Tree)
}
```

```{r}
MSE.T <- c()

for (i in stocknames)
{x = filter(DJI1Training, stock == i)

  y = filter(DJI1Test, stock == i)

  MSE.T <- c(MSE.T,StockFunctionTree(x, y, formulafinal))
  
}

view(MSE.T)

```

### Creating the SVR Model and MSE

```{r}
StockFunctionSVR <- function(training, testing, formula) {
  set.seed(111)
  
  svmtune <- tune.svm(formula, data = training, gamma =seq(.01, 0.1, by = .01), cost = seq(0.1,1, by = 0.1))
  
  SVRModel <- svm(formula, data = training, kernel = "radial", cost = svmtune$best.parameters$cost, gamma = svmtune$best.parameters$gamma)
 
  SVRPreds <- predict(SVRModel, newdata = testing)
  
  MSE.SVR <- (mean((testing$percent_change_next_weeks_price - SVRPreds)^2))

  print(MSE.SVR)
}
```

```{r}
MSE.SVR <- c()

for (i in stocknames)
{x = filter(DJI1Training, stock == i)

  y = filter(DJI1Test, stock == i)

  MSE.SVR <- c(MSE.SVR,StockFunctionSVR(x, y, formulafinal))
  
}

view(MSE.SVR)
```

### Calculating MSE Averages

```{r}
MSEDataFrame <- (cbind(df, MSE.T, MSE.SVR))
MSEDataFrame <- data.frame(MSEDataFrame)
```

```{r}
MSE.LAvg <- mean(MSEDataFrame$MSE.L)
MSE.TAvg <- mean(MSEDataFrame$MSE.T)
MSE.SVRAvg <- mean(MSEDataFrame$MSE.SVR)
```

```{r}
MSE.LAvg
```

```{r}
MSE.TAvg
```

```{r}
MSE.SVRAvg
```

### SVR Model with all variables, for comparison

```{r}
SVRALL <- c()

for (i in stocknames)
{x = filter(DJI1Training, stock == i)

  y = filter(DJI1Test, stock == i)

  SVRALL <- c(SVRALL,StockFunctionSVR(x, y, percent_change_next_weeks_price ~ Week + open + high + low + close + percent_change_price + percent_change_volume_over_last_wk + open.lag + close.lag + high.lag + low.lag + percent_change_volume_over_last_wk.lag))
  
}

view(SVRALL)
```

```{r}
SVRComparison <- cbind(SVRALL, MSE.SVR)
mean(SVRALL)
```

### Creating the CAPM Model and Beta Values

```{r data10}
DJICAPM1 <-split(DJI1Test, DJI1Test$stock)
SP500Data <-read.csv("S&P_500.csv", header = TRUE, sep = ",")
SP500Data1 = na.omit(SP500Data)
SP500Data1$Date <- as.Date(SP500Data1$Date, "%m/%d/%Y")
```

```{r data11}
ReturnSP500 = na.omit(Delt(SP500Data1[,5]))
ReturnAA = na.omit(Delt(DJICAPM1$AA[,7]))
ReturnAXP = na.omit(Delt(DJICAPM1$AXP[,7]))
ReturnBA = na.omit(Delt(DJICAPM1$BA[,7]))
ReturnBAC = na.omit(Delt(DJICAPM1$BAC[,7]))
ReturnCAT = na.omit(Delt(DJICAPM1$CAT[,7]))
ReturnCSCO = na.omit(Delt(DJICAPM1$CSCO[,7]))
ReturnCVX = na.omit(Delt(DJICAPM1$CVX[,7]))
ReturnDD = na.omit(Delt(DJICAPM1$DD[,7]))
ReturnDIS = na.omit(Delt(DJICAPM1$DIS[,7]))
ReturnGE = na.omit(Delt(DJICAPM1$GE[,7]))
ReturnHD = na.omit(Delt(DJICAPM1$HD[,7]))
ReturnHPQ = na.omit(Delt(DJICAPM1$HPQ[,7]))
ReturnIBM = na.omit(Delt(DJICAPM1$IBM[,7]))
ReturnINTC = na.omit(Delt(DJICAPM1$INTC[,7]))
ReturnJNJ = na.omit(Delt(DJICAPM1$JNJ[,7]))
ReturnJPM = na.omit(Delt(DJICAPM1$JPM[,7]))
ReturnKO = na.omit(Delt(DJICAPM1$KO[,7]))
ReturnKRFT = na.omit(Delt(DJICAPM1$KRFT[,7]))
ReturnMCD = na.omit(Delt(DJICAPM1$MCD[,7]))
ReturnMMM = na.omit(Delt(DJICAPM1$MMM[,7]))
ReturnMRK = na.omit(Delt(DJICAPM1$MRK[,7]))
ReturnMSFT = na.omit(Delt(DJICAPM1$MSFT[,7]))
ReturnPFE = na.omit(Delt(DJICAPM1$PFE[,7]))
ReturnPG = na.omit(Delt(DJICAPM1$PG[,7]))
ReturnT = na.omit(Delt(DJICAPM1$T[,7]))
ReturnTRV = na.omit(Delt(DJICAPM1$TRV[,7]))
ReturnUTX = na.omit(Delt(DJICAPM1$UTX[,7]))
ReturnVZ = na.omit(Delt(DJICAPM1$VZ[,7]))
ReturnWMT = na.omit(Delt(DJICAPM1$WMT[,7]))
ReturnXOM = na.omit(Delt(DJICAPM1$XOM[,7]))
```

```{r data12}
DataCAPM = cbind(ReturnSP500, ReturnAA, ReturnAXP, ReturnBA, ReturnBAC, ReturnCAT, ReturnCSCO, ReturnCVX, ReturnDD,
               ReturnDIS, ReturnGE, ReturnHD, ReturnHPQ, ReturnIBM, ReturnINTC, ReturnJNJ, ReturnJPM, ReturnKO,
               ReturnKRFT, ReturnMCD, ReturnMMM, ReturnMRK, ReturnMSFT, ReturnPFE, ReturnPG, ReturnT, ReturnTRV,
               ReturnUTX, ReturnVZ, ReturnWMT, ReturnXOM)
```

```{r data13}
colnames(DataCAPM) = c("SP500", "Alcoa", "American Express", "Boeing","Bank of America","Caterpillar", "Cisco",
                     "Chevron","DuPont","Walt Disney", "General Electric", "Home Depot", "HP", "IBM", "Intel",
                     "J&J", "JPMorgan", "Coca-Cola", "KRFT", "McDonald's", "3M", "Merck", "Microsoft", "Pfizer",
                     "Procter & Gamble", "AT&T", "Travelers", "Raytheon", "Verizon", "Walmart", "Exxon")
head(DataCAPM)
```

```{r, fig.height= 10, fig.width= 25}
boxplot(DataCAPM,main="Expected Return", xlab="Stock Picks", ylab="Return", las = 2)
```

```{r}
DataMean=apply(DataCAPM, 2, mean)
DataSD=apply(DataCAPM, 2, sd)
```

```{r}
cbind(DataMean,DataSD)
```

```{r data14}
DataCAPM1 = data.frame(DataCAPM)
```

```{r data15}
CAPMFunction <- function (stockname)  {
  set.seed(111)

  LinearModel <- lm(stockname ~ SP500, data = DataCAPM1)

  LinearSumm <- (summary(LinearModel))
  
  BetaVal <- (LinearSumm)$coefficients[2, 1]

  print(BetaVal)
}
```

```{r data16}
CAPMNames <- colnames(DataCAPM1[,2:31])
CAPMNames
```

```{r data17}
BetaVals <- c()

for(i in CAPMNames)
{ 
data = DataCAPM1[[i]]  
  BetaVals <- c(BetaVals,CAPMFunction((data)))
  
  
  
}

view(BetaVals)
```

```{r data18}
VariableNames<- c("Alcoa", "American Express", "Boeing","Bank of America","Caterpillar", "Cisco",
                     "Chevron","DuPont","Walt Disney", "General Electric", "Home Depot", "HP", "IBM", "Intel",
                     "J&J", "JPMorgan", "Coca-Cola", "KRFT", "McDonald's", "3M", "Merck", "Microsoft", "Pfizer",
                     "Procter & Gamble", "AT&T", "Travelers", "Raytheon", "Verizon", "Walmart", "Exxon")

BetaVals <- cbind(VariableNames, BetaVals)
BetaVals <- (data.frame(BetaVals))
```

```{r}
BetaVals
```

```{r data19}
BetaVals$BetaVals <- round(as.numeric(BetaVals$BetaVals),4)
BetaVals
```

### Predicting using the SVR Model

```{r data20}
SVRPredsValues <- function(training, testing, formula) {
  set.seed(111)
  
  svmtune <- tune.svm(formula, data = training, gamma =seq(.01, 0.1, by = .01), cost = seq(0.1,1, by = 0.1))
  
  SVRModel <- svm(formula, data = training, kernel = "radial", cost = svmtune$best.parameters$cost, gamma = svmtune$best.parameters$gamma)
 
  SVRPreds <- predict(SVRModel, newdata = testing)

  print(SVRPreds)
}
```

```{r data21}
SVRNextWeek <- c()

for (i in stocknames)
{
x = filter(DJI1Training, stock == i)

y = filter(DJI1Test, stock == i & Week == 25)

SVRNextWeek <- c(SVRNextWeek,SVRPredsValues(x, y, formulafinal))
}

view(SVRNextWeek)
```

### Combining Values

```{r data22}
SVRNextWeek1 <- data.frame(SVRNextWeek)
SVRNextWeek1 <- cbind(VariableNames,SVRNextWeek1)

```

```{r data23}
ComparisonDF <- cbind(SVRNextWeek1, BetaVals$BetaVals)
ComparisonDF <- ComparisonDF %>% 
  rename("Percent Change Prediction for Next Week"  = SVRNextWeek,
         "Beta Value" = "BetaVals$BetaVals")
```

```{r Final}
head(ComparisonDF,30)
```